Capon Beamforming Applied on Moving Objects - An Analysis of Local Lateral Shift-Invariance

### 2013-12-01 HRUIofMovingObjects/jrnl_article.tex
writing

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object which is sensitive to local spatial shifts, the system is said to be spatially locally shift-invariant. This can for instance happen if the axial or lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. In previous literature on Capon beamforming for medical ultrasound imaging however only single frame scenarios have been simulated, and temporal behaviour and effects caused by the increased resolution and lack of oversampling have therefore been neglected.

In this paper we investigate the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We then discuss and investigate different methods in order to improve on it. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer locally laterally shift-variant. We then show that the local lateral shift-invariance can be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer for ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The adaptive nature of the method has been shown to greatly improve the lateral resolution in ultrasound images. In recent work we introduced a fast implementation of the Capon beamformer, capable of processing a stream of images at interactive frame rates \cite{Asen}. It is therefore an interesting question how a moving object appears after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

@@

As mentioned, Capon beamforming has been shown to provide improved lateral resolution and edge definitions in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplishes this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets (SVUS) \cite{Lorenz2005, Rubsamen2013}. The first two, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics \cite{Synnevag2007a}. SVUS reduces the signal-nulling effect of the Capon beamformer which appears if the signal is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the method obtains its super resolution capabilities, but the signal can also get lost in the process. In this paper we make use of a steering vector set to reduce this self-nulling effect.

@@

In order to visualize the local lateral shift-invariance of a given imaging system, we will make use of lateral shift-variance plots (LSV-plots), introduced by Hergum \textit{et al.} \cite{Hergum2007}. A LSV-plot is constructed for a given imaging system by imaging a point scatterer moving laterally at constant speed. For each lateral position the root-mean-square (RMS) of the image data is calculated in range, and the resulting beam profiles are stacked in to an image. For a given scan sequence there are twice as many point positions than there are receive lines. The LSV-plots in this paper are displayed as a contour plot with contours on -1, -2, -3, -6, -12 and -24 dB. The maximum for each beam profile is also marked with a black dot, and the amplitude variation among these points is presented in a subplot next to the LSV-plot.

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). The red dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse with a $\text{f-number}=3$ and corresponding cosine apodization on receive. The transmit focus in azimuth and elevation was configured to be equal and the point scatterer was moved through this focus. The lateral sampling density has been doubled prior to detection through coherent interpolation. We see that the contours are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. In this, and all the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located exactly on this beam at one time instance.

The scalloping loss in Fig.\,\ref{fig:das_shiftinvariant} is smaller than what was predicted by Fig.\,\ref{fig:das_capon_beams}. To understand why note that Fig.\,\ref{fig:das_capon_beams} shows the scanned response from a narrowband system while the imaging system in Fig.\,\ref{fig:das_shiftinvariant} uses a broadband pulse. The beam in Fig.\,\ref{fig:das_shiftinvariant} while therefore be more similar to a squared gaussian function than the squared $sinc$ functions present in Fig.\,\ref{fig:das_capon_beams}. The imaging system in Fig.\,\ref{fig:das_shiftinvariant} also uses an expanding cosine window on receive.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2a.eps}\label{fig:das_shiftinvariant}}
\hfill{}
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2b.eps}\label{fig:das_shiftvariant}}
}
\caption{LSV-plots of delay-and-sum beamforming. If the contours are diagonal and the amplitude variation among the peak points is small, the imaging system is said to be locally lateral shift-invariant. Red vertical lines are plotted where transmit-receive beam pairs are located. a) Sampling corresponding to (\ref{eq:resolution}). The system is locally lateral shift-invariant. b) Undersampling with a factor of two. The system is no longer locally lateral shift-invariant.}
\label{fig:das}
\end{figure*}

@@

\subsection{Oversampling on Transmit}
The straight forward way of obtaining sufficient sampling in an active system is by increasing the number of transmit beams (tx-beams) until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
}
\caption{Capon beamforming as described in Fig\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed compared with Fig\,\ref{fig:capon} to better see the beam-to-beam variation. a) 4x oversampling. b) 10x oversampling.}
\label{fig:capon_oversampling}
\end{figure*}

@@

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than ten times oversampling is required in order to achieve lateral shift invariant imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by only using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length (coarse-fine beamforming \cite{Thomenius}). It is therefore possible to make a sweep between a given beam and its two nearest neighbours using only phase rotation without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

For an uniform linear array located along the x-axis the steering vector $\vec{a}$ should be calculated as 
\begin{align}
\vec{a} = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swiped from $-\Delta/2$ to $\Delta/2$ where $\Delta$ is as defined in (\ref{eq:resolution}). Note that the steering vector is calculated for a $L$ element array centered around the origin, even though the full array has $M$ elements. It is important that this subarray has the same element pitch as the full array.

@@

@@

\section{Results}\label{sec:res}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in-vitro} data with and without oversampling by phase rotation. The data was acquired using a modified version of the Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S transducer in harmonic mode. The transmit beam density was set to match (\ref{eq:resolution}). A tissue mimicking phantom containing wire targets and cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:image_das} shows the first frame out of 14 in the acquired loop. Movies of the whole loop processed with DAS and Capon with and without 10 times oversampling by phase rotation can be found in the supplementary material.

In Fig\,\ref{fig:image_das} one point scatterer has been selected (red box), and then used to generate LSV-plots. Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using DAS beamforming, Fig\,\ref{fig:phantom_capon_LSV} shows the same for the Capon beamformer without oversampling, and Fig\,\ref{fig:phantom_capon_LSV_10PR} presents the result of applying 10 times oversampling by phase rotation. As before, vertical dotted lines indicates rx-tx beam pairs.

@@

\begin{figure}\caption{Plot of contrast vs oversampling factor (phase rotation) in in-vivo data.}
\end{figure}

@@

\begin{figure*}[!t]
\centerline{
\begin{tabular}{cc}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img LSV_image.eps}\label{fig:image_das}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig7b.eps}\label{fig:phantom_das}}\\
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig7c.eps}\label{fig:phantom_capon_LSV}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig7d.eps}\label{fig:phantom_capon_LSV_10PR}}
\end{tabular}
}
\caption{LSV-plots generated from \textit{in-vitro} phantom data. The probe was manually dragged across the phantom. See the supplementary material for videos of the whole loop \comment{Add ref to videos.} a) First frame in the recorded loop. A red box indicates the point scatterer used to construct the LSV-plots.  b) Delay-and-sum beamforming. The amplitude variation of the point scatterer is minimal. c) Capon beamforming without oversampling. In the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we get a high amplitude value. In all the other frames the signal loss is massive. d) Capon with 10 times oversampling using phase rotation. The amplitude of the point scatterer is better preserved in all frames. Note the the oscillation along the diagonal does not indicate shift-variance. It is a result of having to low frame rate and moving the probe to fast.}
\label{fig:LSV_invitro}
\end{figure*}

@@

\section{Discussion}\label{sec:dis}
Discuss results presented in Section \ref{sec:res}.

When signal detection (absolute value and square root) is performed before an ultrasound image is displayed the lateral bandwidth is increased. In order to not introduce aliasing in the image, IQ-interpolation is typically used on ultrasound systems to double the lateral beam density before detection. For the same reason we need to increase the lateral sampling when Capon beamforming is applied, but now on a channel data level. The Capon beamformer improves the lateral resolution, hence the lateral bandwidth is increased, and higher lateral sampling is required.

@@

The figure below shows a lateral shift invariance plots for a well sampled delay-and-sum image (a), the same sampling with Capon processing (b), and 4x over sampling processed with Capon. The shift invariance plots the beamprofile (TODO: Make the shift invariance plot). The results show that when using the Capon beamformer we have to over sample with a factor four to remove the blinking artifact. This can be done without affecting frame rate by using the available parallel receive-lines found on most ultrasound systems. May also use phase rotation.

Other methods claiming super resolution would also need oversampling (Eigenspace, LCA etc.) to some extent.

@@

One way of reducing the number of matrix inversions in PRB is to estimate the covariance matrix from multiple beams \cite{Jensen2012}, generating one covariance matrix per range.  Tailoring the method of Jensen and Austeng to cardiac ultrasound imaging it makes more sense, due to motion, to construct one covariance matrix per tx-beam using multiple parallel rx-beams. With this solution the number of matrix inversions will not changes, but the time taken to construct the matrices will increase. We will not investigate this method further in this paper. \comment{Kanskje litt dumt å nevne det da...?} \comment{move to discussion}

The phase rotation method in this paper is similar to a SVUS, but where SVUS typically has a criteria for selecting the best output after applying all the steering vectors we output all of them.

Oversampling by phase rotation introduce similar block artifacts as PRB, but on a smaller scale. These artifacts could be removed if we distributes our steering angles as beams are spread out for the STB technique. We then get a phase rotation version of STB where e.g. eight phase rotated beams are averaged into four. \comment{Move to discussion?}

Phase rotation will introduced block artifacts equal to the once observed when using PRB. This can be improved on by implementing a phase rotation version of STB where 8PR are converted to 4PR-STB by averaging 4 phase rotated outputs from two beams rotated to the same angle.

The geometric distortion remaining in Fig. 7d should be possible to fix by a phase rotation version of STB.

\subsection{Impact on Real Time Performance}
Discuss, and show with benchmarks how the different methods for making Capon beamforming shift-invariant impact the total execution time.

\section{Conclusion}\label{sec:con}
...

>>>>

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object which is sensitive to local spatial shifts, the system is said to be spatially locally shift-variant. This can for instance happen if the axial or lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. In previous literature on Capon beamforming for medical ultrasound imaging however only single frame scenarios have been simulated, and temporal behaviour and effects caused by the increased resolution and lack of oversampling have therefore been neglected.

In this paper we analyse the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We then discuss and investigate different methods in order to improve on it. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer locally laterally shift-variant. We then show that the local lateral shift-invariance can be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer (also known as the minimum variance beamformer) for ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The adaptive nature of the method has been shown to greatly improve the lateral resolution in ultrasound images. In recent work we introduced a fast implementation of the Capon beamformer, capable of processing a stream of images at interactive frame rates \cite{Asen}. It is therefore an interesting question how a moving object appears after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

@@

As mentioned, Capon beamforming has been shown to provide improved lateral resolution and edge definitions in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplishes this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets (SVUS) \cite{Lorenz2005, Rubsamen2013}. The first two, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution settings \cite{Synnevag2007a}. SVUS reduces the signal-nulling effect of the Capon beamformer which appears if the signal is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the method obtains its super resolution capabilities, but the signal can also get lost in the process. In this paper we make use of a steering vector set to reduce this self-nulling effect.

@@

In order to visualize the local lateral shift-invariance of a given imaging system, we will make use of lateral shift-variance plots (LSV-plots), introduced by Hergum \textit{et al.} \cite{Hergum2007}. A LSV-plot is constructed for a given imaging system by imaging a point scatterer moving laterally at constant speed. For each lateral position the root-mean-square (RMS) of the image data is calculated in range, and the resulting beam profiles are stacked in to an image. For a given scan sequence there are twice as many point positions than there are receive lines. The LSV-plots in this paper are displayed as a contour plot with contours on -1, -2, -3, -6, -12 and -24 dB. The maximum for each beam profile is also marked with a black dot, and the amplitude variation among these points is presented in a subplot next to the LSV-plot. \comment{Add interp factor}

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). The red dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse with a $\text{f-number}=3$ and corresponding cosine apodization on receive. The transmit focus in azimuth and elevation was configured to be equal and the point scatterer was moved through this focus. The lateral sampling density has been doubled prior to detection through coherent interpolation. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. In this, and all the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located exactly on this beam at one time instance.

The scalloping loss in Fig.\,\ref{fig:das_shiftinvariant} is smaller than what was predicted by Fig.\,\ref{fig:das_capon_beams}. To understand why note that Fig.\,\ref{fig:das_capon_beams} shows the scanned response for a narrowband system while the imaging system in Fig.\,\ref{fig:das_shiftinvariant} uses a broadband pulse. The beam in Fig.\,\ref{fig:das_shiftinvariant} while therefore be more similar to a squared gaussian function than the squared $sinc$ functions depicted in Fig.\,\ref{fig:das_capon_beams}. The imaging system in Fig.\,\ref{fig:das_shiftinvariant} also uses an expanding cosine window on receive.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2a.eps}\label{fig:das_shiftinvariant}}
\hfill{}
\subfloat[]{\includegraphics[width=0.5\linewidth]{\img fig2b.eps}\label{fig:das_shiftvariant}}
}
\caption{LSV-plots of delay-and-sum beamforming. If the contours are diagonal and the amplitude variation among the peak points (black dots) is small, the imaging system is said to be locally laterally shift-invariant. Red vertical lines are plotted where transmit-receive beam pairs are located. a) Sampling corresponding to (\ref{eq:resolution}). The system is locally laterally shift-invariant. b) Undersampling with a factor of two. The system is no longer locally laterally shift-invariant.}
\label{fig:das}
\end{figure*}

@@

\subsection{Oversampling on Transmit}
The straight forward way of obtaining sufficient sampling in an active system is by increasing the number of transmit beams (tx-beams) until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten \comment{Increase the sampling to 16x to get the scalloping loss below 1 dB?} respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
}
\caption{Capon beamforming as described in Fig\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed compared with Fig\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling. b) Ten times oversampling.}
\label{fig:capon_oversampling}
\end{figure*}

@@

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than ten times oversampling is required in order to achieve lateral shift invariant high resolution imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by only using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length (See coarse-fine beamforming in \cite{Thomenius}). It is therefore possible to make a sweep between a given beam and its two nearest neighbours using only phase rotation without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

For an uniform linear array located along the x-axis the steering vector $\vec{a}_\theta$ should be calculated as 
\begin{align}
\vec{a}_\theta = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swiped from $-\Delta/2$ to $\Delta/2$ where $\Delta$ is as defined in (\ref{eq:resolution}). Note that the steering vector is calculated for a $L$ element array centered around the origin, even though the full array has $M$ elements. It is important that this subarray has the same element pitch, $x_i - x_{i-1}$, as the full array.

@@

Since the phase rotation method operates on the same data as before, there will be no reduction in the acquisition frame rate by applying this method. In addition, since the same number of covariance matrices are constructed and inverted, there is no extra cost involved here as well. The only additional cost will be the multiplication of the extra steering vectors with the inverted covariance matrix. If the equation $\mat{\hat{R}}\vec{a} = \vec{b}$, as in \cite{Asen}, is solved using Gauss-Jordan elimination, bringing matrix $[\mat{\hat{R}}  \vec{a}_{\theta_1} \vec{a}_{\theta_2} \cdots \vec{a}_{\theta_{N_\theta}}]$ into a reduced row echelon form has the same complexity as for inverting $\mat{\hat{R}}$, as long as $N_\theta < L$.

@@

\section{In-vitro phantom data}\label{sec:res}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in-vitro} data with and without oversampling by phase rotation. The data was acquired using a modified Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S transducer operating in harmonic mode. The transmit beam density was set to match (\ref{eq:resolution}). A tissue mimicking phantom containing wire targets and cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:image_das} shows the first frame out of 14 in the acquired loop. %Movies of the whole loop processed with DAS and Capon beamforming with and without ten times oversampling by phase rotation can be found in the supplementary material \multimedia{Videos}.

In Fig\,\ref{fig:image_das} one point scatterer has been selected (red box \comment{convert into a thicker white box}), and then used to generate LSV-plots.  \comment{Todo: Add lateral coherent interpolation as done for the simulations? As for now speckle is not fully resolved without it.}

Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using the DAS beamformer. The scalloping loss is less than 1 dB, and the system is therefore locally laterally shift-invariant. As a consequence point scatterers has a consistent appearance when subjected to motion \multimedia{Video DAS}.

Figure \ref{fig:phantom_capon_LSV} shows the same for the Capon beamformer without oversampling. In the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we get a high amplitude value at this location. In almost all the other frames the signal loss is massive. The result is blinking point scatterers when subjected to motion \multimedia{Video Capon}.

Figure \ref{fig:phantom_capon_LSV_10PR} presents the result of applying ten times oversampling by phase rotation. We observe how the amplitude of the point scatterer is better preserved in all frames, and the scalloping loss is not far from our 1 dB threshold. Note that the oscillation along the diagonal does not indicate shift-variance. It is a result of lack of control when the probe was manually dragged across the phantom. In order to get a continues straight diagonal the probe has to be dragged across the phantom with a speed which match the acquisition frame rate. In addition, due to the increase in lateral resolution the Capon beamformer requires smaller frame-to-frame movements than DAS. Looking at the video of Capon beamforming with ten times oversampling by phase rotation \multimedia{Capon + 10PR} there is no visible blinking anymore. We therefore conclude that the system is now locally laterally shift invariant.

@@

%\begin{figure}\caption{Plot of contrast vs oversampling factor (phase rotation) in in-vivo data.}
%\end{figure}

@@

\begin{figure*}[!t]
\centerline{
\begin{tabular}{cc}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img LSV_image.eps}\label{fig:image_das}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig7b.eps}\label{fig:phantom_das}}\\
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig7c.eps}\label{fig:phantom_capon_LSV}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig7d.eps}\label{fig:phantom_capon_LSV_10PR}}
\end{tabular}
}
\caption{LSV-plots generated from \textit{in-vitro} phantom data where the probe has been manually dragged across the phantom. See the supplementary material for videos of the whole loop \multimedia{Videos}. Capon beamforming has been performed with the same settings as in Fig.\,\ref{fig:capon1}. a) First frame in the recorded loop. A red box indicates the point scatterer used to construct the LSV-plots.  b) Delay-and-sum beamforming. c) Capon beamforming without oversampling. d) Capon with ten times oversampling using phase rotation.}
\label{fig:LSV_invitro}
\end{figure*}

@@

\section{Discussion}\label{sec:dis}
In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. In Section \ref{} it was shown that a Capon beamformer with standard settings is not locally laterally shift-invariant when subjected to motion. In Section \ref{} and \ref{} we presented restored

%Discuss results presented in Section \ref{sec:res}.

%When signal detection (absolute value and square root) is performed before an ultrasound image is displayed the lateral bandwidth is increased. In order to not introduce aliasing in the image, IQ-interpolation is typically used on ultrasound systems to double the lateral beam density before detection. For the same reason we need to increase the lateral sampling when Capon beamforming is applied, but now on a channel data level. The Capon beamformer improves the lateral resolution, hence the lateral bandwidth is increased, and higher lateral sampling is required.

@@

%The figure below shows a lateral shift invariance plots for a well sampled delay-and-sum image (a), the same sampling with Capon processing (b), and 4x over sampling processed with Capon. The shift invariance plots the beamprofile (TODO: Make the shift invariance plot). The results show that when using the Capon beamformer we have to over sample with a factor four to remove the blinking artifact. This can be done without affecting frame rate by using the available parallel receive-lines found on most ultrasound systems. May also use phase rotation.

%Other methods claiming super resolution would also need oversampling (Eigenspace, LCA etc.) to some extent.

@@

%One way of reducing the number of matrix inversions in PRB is to estimate the covariance matrix from multiple beams \cite{Jensen2012}, generating one covariance matrix per range.  Tailoring the method of Jensen and Austeng to cardiac ultrasound imaging it makes more sense, due to motion, to construct one covariance matrix per tx-beam using multiple parallel rx-beams. With this solution the number of matrix inversions will not changes, but the time taken to construct the matrices will increase. We will not investigate this method further in this paper. \comment{Kanskje litt dumt å nevne det da...?} \comment{move to discussion}

%The phase rotation method in this paper is similar to a SVUS, but where SVUS typically has a criteria for selecting the best output after applying all the steering vectors we output all of them.

%Oversampling by phase rotation introduce similar block artifacts as PRB, but on a smaller scale. These artifacts could be removed if we distributes our steering angles as beams are spread out for the STB technique. We then get a phase rotation version of STB where e.g. eight phase rotated beams are averaged into four. \comment{Move to discussion?}

%Phase rotation will introduced block artifacts equal to the once observed when using PRB. This can be improved on by implementing a phase rotation version of STB where 8PR are converted to 4PR-STB by averaging 4 phase rotated outputs from two beams rotated to the same angle.

%The geometric distortion remaining in Fig. 7d should be possible to fix by a phase rotation version of STB.

%Empirical testing is required to find how large the oversampling needs to be. Typically it depends on how bright the point scatterer or edge are with respect to the background. Cox \cite{Cox1973} derives formula for how large the oversampling has to be for a given signal-to-noise scenario. It is therefore possible to measure the highest bright-point-to-speckle value and derive the level of oversampling from this value.

%\subsection{Impact on Real Time Performance}
%Discuss, and show with benchmarks how the different methods for making Capon beamforming shift-invariant impact the total execution time.

\section{Conclusion}\label{sec:con}
In this paper we have investigated the lateral shift-invariance properties of the Capon beamformer when imaging moving objects. It was shown that Capon beamforming with the same beam density as required for delay-and-sum beforming results in blinking point scatterers when subjected to motion.  We then discussed and investigated different methods in order to reduced this effect. It was then shown that the local lateral shift-invariance could be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.
